\subsection{Causal Discovery Methods}

In this section we give an impression of the diversity of causal discovery, by describing some well-known methods in more detail. We give a concise description of the algorithm and the assumptions that it relies on.

\subsubsection{Constraint-Based Causal Discovery}
Constraint-based approaches derive constraints from the data, and use it to restrict the class of possible underlying causal graphs. Causal inference is treated as a constraint-satisfaction problem. Generally, CIRs are chosen as constraints, which can be used to derive the separation of variables and the orientation of edges in the graph. We first describe three global methods IC, PC and FCI, and then three local methods Y-Structures, LCD and ICP.

\textbf{Inductive Causation} (IC) was introduced by \citet{verma1991equivalence} and generally describes how we can induce a Partially Directed Acyclic Graph\footnote{A PDAG is a graph without directed cycles, edges may be directed or undirected} (PDAG) from conditional independences in the data. The algorithm follows two steps: inducing the skeleton and orienting the edges. For every pair of variables $X$ and $Y$, we check whether there is an edge in the PDAG. Using the faithfulness assumption, we add an edge if there is no separating set of variables $\B{S}_{XY}$ that makes $X$ and $Y$ conditionally independent ($\nexists \B{S}_{XY} : \indep{X}{Y}{\B{S}_{XY}}{}$). One easy step of edge orientation uses the separation criterion of colliders. If two non-adjacent variables $X$ and $Y$ share a neighbour $Z$ that is not in $\B{S}_{XY}$, it must be a collider on the path and we induce edge orientation $X\to Z \ot Y$. Application of additional orientation rules lead us to a maximally oriented PDAG, which describes the MEC of graphs that induce the joint data distribution. One early set of such rules was described by \citet{spirtes2000causation} in the SGS algorithm, which was named after the authors.
            
IC is quite limited, because it relies on several assumptions about the underlying SCM (e.g. causal sufficiency), and its naive implementation is costly due to the search over all separating sets $\B{S}_{XY}$.

\textbf{PC}, named after its inventers Peter Spirtes and Clark Glymour \citep{spirtes1991algorithm}, reduces the cost of naive IC. A systematic algorithm finds the separating sets $\B{S}_{XY}$ in polynomial time. Starting from a fully-connected graph, edges are systematically removed by considering separating sets of increasing cardinality, and only taking into account the variables that neighbour $X$ and $Y$. For example, first edges are removed between variable pairs that are independent given the empty set (cardinality 0). Then, edges are removed between remaining adjacent variable pairs that are independent given one of their neighbours. Already some possible separating sets can be skipped here, because edges were removed in the previous step. Therefore, as we consider larger possible separating sets, the number of neighbours to choose from decreases.

\textbf{Fast Causal Inference} (FCI) extends PC to allow for selection bias and latent confounding. Dropping the causal sufficiency assumption means that it should consider some possible separating sets $\B{S}_{XY}$ that contain variables not adjacent to $X$ or $Y$ (specifically, their ancestors). FCI is a feasible algorithm for datasets with many variables when the underlying graph is sparse and bidirected edges (i.e. confounded variables) are not too much chained together. It was first introduced by \citet{spirtes1999algorithm}, and gradually developed since then. A modern version named FCI+ by \citet{claassen2013learning} is relatively fast. It finds the Partially Ancestral Graph (PAG)\footnote{The PAG is a graph that can represent latent confounding.} in polynomial time.
          
\textbf{Y-structures} are a pattern in a PAG, which has information about ancestral relations between four random variables. \citet{mooij2015empirical} showed that a set of independence tests can be used to infer some ancestral relations from observational data. This local method can be used to find an incomplete set of ancestral relations in the underlying SCM. 

Specifically, we marginalize over four random variables $X$, $Y$, $Z$ and $U$. Then we perform four statistical tests: $\indep{Z}{Y}{X}{}$, $Z \nCI Y$, $\dep{Z}{U}{X}{}$ and $Z \CI U$. If all tests are positive, there are only two possible PAGs representing the marginalization over the four variables, which are called Y-structure (a term coined for a score-based method by \citet{mani2006bayesian}) and Extended Y-structure. In both PAGs, we can use the backdoor criterion \citep{pearl2009causality} to infer that $p(Y\given \intervene(X=x))=p(Y\given X)$.

\citet{mooij2015empirical} investigate the performance effect of adding more independence tests. By adding two more tests, the Y-structure remains the only possible PAG. After that, more redundant independence tests can be added. Adding more tests necessarily reduces recall, but precision can improve if the extra tests eliminate false positives. Interestingly, in their experiments on synthetic data, maximum precision is not always achieved with the minimum or maximum number of tests.

An interesting insight from \citet{mooij2015empirical} is that the faithfulness assumption becomes more problematic as the number of random variables grows. The data is sampled from a SCM, which makes the faithfulness assumption very reasonable. However, it appears that the marginalized data can be almost faithless to the supposed PAG. 

\textbf{Local Causal Discovery} (LCD) is a local method that looks at ancestral causal relations in the context of a changing environment. In the data, we marginalize over three variables ($C, X, Y$), one of which is exogenous ($C$). Next, we perform three statistical tests: $\indep{C}{Y}{X}{}$, $C \nCI X$ and $X \nCI Y$. \citet{cooper1997simple} proves that there is a relationship $X \to Y$ if these tests are positive.

This method allows for latent confounders, and assumes that there is no selection bias. Note that this method can find only a subset of ancestral relations. The method is proven by listing all possible causal graphs of three variables in which the context variable is not caused by endogenous variables. Out of the 32 networks in which $X$ causes $Y$, LCD is only able to identify this relation in 3 cases. The others are not identified, because there are other configurations that induce the same independence relations. For example, if $C$ causes $Y$ not only through $X$, but via another path as well, we will not find that $X$ causes $Y$.  

% e.g. Trigger?\citep{chen2007harnessing}. Local strategy.

\textbf{Invariant Causal Prediction} (ICP) is a more general local method that also uses the context of a changing environment. In its original formulation by \citet{peters2016causal} it outputs a subset of parents of a target variable, assuming causal sufficiency and acyclicity. Faithfulness is not required. \citet{mooij2016joint} show that when faithfulness is assumed, we no longer require the causal sufficiency and acyclicity assumptions, and the algorithm outputs ancestors.

When we model the context (e.g. experimental setting) as a variable that is not a direct cause of the target variable, then the conditional distribution of the target variable given its direct causes is invariant to the value of the context variable. This property is used by ICP to find parent or ancestor relations in data from different contexts. 

In a naive implementation, we would have to investigate every possible parent set for every target variable, making the complexity exponential in the number of variables. Since (direct) causes tend to correlate with their effect, it is reasonable to only consider variables that have a relatively high correlation with the target variable. Such an approximation was used by \citet{peters2016causal} and \citet{meinshausen2016methods} to apply ICP to the dataset of \citet{kemmeren2014large} with over six thousand variables.

    
\subsubsection{Score-Based Causal Discovery}

Score-based approaches provide an alternative to the methods discussed above. They search for the causal graph that optimizes some loss function, often based on independences in the data. Two global methods are discussed below.
% (e.g., Cooper and Herskovits,1992; Heckerman et al., 1995; Chickering, 2002; Koivisto and Sood, 2004) (see: Peters 7.2)
% Mani et al 2006: disadvantages of constraint-based (citing Heckerman et al., 1999): (1) the inability to include prior belief in the form of structure probabilities, (2) the output is based on the significance level used for the independence tests, and (3) there is no quantification of the strength of the hypothesis that is output.

\textbf{Accounting for Strong Dependences} (ASD) is a method that focusses on graphs satisfying dependences in the data, whereas most methods focus on the independences. The groundwork of the method was done by \citet{hyttinen2014constraint}. Dependence and independence relations in the data are encoded as soft constraints in ASP (Answer Set Programming), together with rules that determine that the solution should be a valid graph. The solution is then found by an ASP solver, that minimizes the loss computed as the sum of the weights of the constraints that are not satisfied. This method can only be applied to problems with a small number of variables, because the number of dependence and independence relations quickly becomes very large.

\citet{magliacane2016ancestral} compute weights based on the p-value of the conditional independence test and the significance level, such that strong dependences obtain a larger weight than independences. They also provide a method to compute a confidence score for single features of the graph (like an ancestral relation) by solving two optimization problems and subtracting the losses. In one problem, the presence of the feature is added as a hard constraint, in the other the absence is added as a hard constraint. 

As an implementation of their Joint Causal Inference (JCI) framework, \citet{mooij2016joint} adapt the ASD method to include interventional data. In their adaptation, some constraints are added to the ASP problem that encode the JCI assumptions for interventional data.

\textbf{Greedy Equivalence Search} (GES) was first introduced by \citet{meek1997graphical} and further detailed by \citet{chickering2002optimal}. A search for the Markov equivalence class (MEC) is performed in two phases. The graph is initialized without edges. In the first phase, we move between MECs by adding edges. In each step, we consider the set of MECs that are one edge addition away from the current MEC. Reversing a covered edge\footnote{An edge $X\to Y$ is covered if $X$ and $Y$ have the same parents (and $Y$ has $X$ as parent as well)} does not change the MEC, so we need to consider every combination of covered edge reversals, followed by a single edge addition, followed again by covered edge reversals. We score the MECs and move to the one with the highest score. \citet{meek1997graphical} proposed the Bayesian scoring criterion to score DAGs in a MEC. Under some conditions this can be approximated by the Bayesian Information Criterion (BIC). 

Once we reach a local maximum, \citet{chickering2002optimal} shows that the MEC contains the generative distribution. The second phase is analogous to the first, but this time single edges are removed. \citet{chickering2002optimal} shows that this final MEC is a perfect map of the generative distribution.

\citet{hauser2012characterization} generalize the method to datasets with interventional data (GIES). They introduce the interventional MEC as the object of the search. The algorithm is adapted by introducing the turning phase. Repeated application of the three phases leads to the interventional MEC underlying the generative distribution. They note that the space of interventional MECs is more fine-grained, meaning that this type of data leads to improved identifiability.


\subsubsection{Hybrid methods}
% \textbf{Max-Min Hill-Climbing} (MMHC) (Tsamardinos et al., 2006), for example, alternates between constraint-based and score-basedupdates.

Finally, there are also methods that rely on both constraints and the optimisation of some score. SP is the most relevant example, and is described below.

\textbf{Sparsest Permutation} (SP) is a global method proposed by \citet{raskutti2018learning} that searches for the MEC in a space of orders of random variables, using observational data. The order is a permutation of variables in which ancestors precede descendents. Given an order and a method to infer dependence and independence relations, we can infer a DAG that satisfies causal minimality. The SP algorithm searches over the DAGs infered from all possible permutations and selects the DAG with the smallest number of edges. \citet{raskutti2018learning} show that this algorithm is consistent, which means that it finds a DAG in the correct MEC as sample size tends to infinity. This consistency relies on an assumption that is a weaker version of faithfulness. 

\citet{solus2017consistency} propose a greedy algorithm (GSP) that increases the number of variables that can be handled from about 10 to hundreds. The only sacrifice is that a somewhat stronger assumption is required, which is still weaker than faithfulness. They introduce the DAG associahedron, a sub-polytope of the permutohedron which indicates possible directions for the search. These directions are based on covered edges. 

\citet{wang2017permutation} adapt the algorithm further to take interventional data into account (IGSP). The score is now the sum over the individual scores for each intervention distribution. This score is a weighted sum of the number of edges in the inferred graph, excluding edges towards the intervened variable, and a maximum likelihood score of the interventional data given some model assumptions.



% first with consistency proof. (permutation-based without proof:  (Bouckaert, 1992; Singh and Valtorta, 1993; Teyssier and Koller, 2005). )


% \subsubsection{MOVE TO CONSTRAINT BASED? Other constraints?}
% other statistical patterns in the joint distribution can be exploited too (e.g Mooij et al., 2016; Peters et al., 2017)

% TODO: IDA (Maathuis et al. 2010), MMHC, CV-Lasso regression?, CI?

% SOURCE?
% Focus here on independence, but there are other patterns that can be exploited, such as 
% * “Verma constraints” (Shpitser et al.,2014), 
% * algebraic constraints in the linear-Gaussian case (van Ommen and Mooij, 2017),
% * non-Gaussianity in linear models (Kano and Shimizu, 2003), and 
% * non-additivity of noise in nonlinear models (Peters et al., 2014) can also be exploited.